\vspace{-2cm} Government websites as data: A methodological pipeline with application to the websites of municipalities in the United States

### 2018-07-20 paper/manuscript.tex
finished revisions to data section

% BRUCE HERE
In this section we introduce the data we use in our application---the analysis of municipal websites in six states - Indiana, Louisiana, New York, Washington, California and Texas. These states provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions is represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities less densely populated areas are covered. Furthermore, the sample is politically balanced -  we have three blue states (CA, WA, NY) and three red states (TX, IN, LA). Finally, our dataset contains some of the wealthiest states (NY, CA, WA and TX are \#2, \#8, \#9 and \#16 respectively, by GDP per capita \citep{BureauofEconomicAnalysis2017}), but also some of the poorer ones (IN and LA). In terms of pure GDP per capita, the sample is on the less affluent side - however, wealth is also correlated with poverty: CA is the state with the highest poverty rate in the country, and LA, NY and TX follow closely \citep{Fox2017}.

We acquired the website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. This method proved to be very reliable. Two, the General Services Administration (GSA) maintains all .gov addresses, and provides a complete\footnote{Domains used for testing and internal programs are excluded.} list of all such domains to the public through GitHub\footnote{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}\footnote{This list is updated once per month - we rely on the version released on January 16, 2017. The data from the GSA contains the following data: One, domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'). Two, the type of government entity to which the domain is registered, such as city, county, federal agency, etc. Three, for federal agencies, the name is specfied. Finally, the city in which the domain is registered, is noted.}. Naturally, this list does not contain cities which do not use a .gov website (or, in many cases, a city owns a registered .gov address, but uses a different one),. Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa).

Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the former, and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

To test whether the websites we found actually work, we use a webdriver-controlled browser (Firefox/Selenium/Geckodriver). This is necessary because a) some city websites simply don't work, and more often, b) cities sometimes change their websites' URLs, in which case they redirect from the old to the new URL. A webdriver-controlled browser, unlike the more rigid conventional scraping tools, will simply follow this redirection. This allows us to subsequently record and use the new URL for the actual website scraping.

The partisanship of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the LEAP project\footnote{\url{http://www.leap-elections.org/}}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and recorded candidates who received money from party committees. For California and Texas, where our data consists of major cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) this method once again proved to be very reliable, and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources.

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments - either in the weak or strong mayor variant - there is a certain degree of variance in where a city's executive authority lies. Unfortunately we do not have access to information about the type of governments across the breadth of our dataset and therefore have to assume that the partisanship of a city's mayor corresponds to the partisanship of the entity responsible for determining the content on its website. Given the prominent place of mayors tend to have on their cities' websites, as well as the fact that council-manager city governments tend to appoint mayors rather than elect them (meaning there are unlikely to be many in our dataset), we feel that any bias arising from this nuance should be minor at the most.

Information on other covariates (population and median household income - from the American Community Survey 5) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.
\input{tables/tabCitiesStatesParties.tex}

@@

For some cities, whose websites make heavy use of JavaScript, this method does not lead to satisfying results. Consequently we restricted our corpus to cities with at least 3 documents.

\section{The Web to Text Pipeline}
In this methodological pipeline from native website files to text data that is appropriate for comparative analysis, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another, but is uninformative regarding policy or process differences between governments. We describe these methodological steps in this section.

>>>>

In this section we introduce the data we use in our application. For data availability reasons, on which we elaborate below, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California and Texas. These states provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions is represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities and less densely populated areas are covered. Furthermore, the sample is politically balanced---we have three blue states (CA, WA, NY) and three red states (TX, IN, LA). Finally, our dataset contains some of the wealthiest states (NY, CA, WA and TX are \#2, \#8, \#9 and \#16 respectively, by GDP per capita \citep{BureauofEconomicAnalysis2017}), but also some of the poorer ones (IN and LA).

We acquired the website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. Two, the General Services Administration (GSA) maintains all .gov addresses, and provides a complete list of all such domains to the public. \footnote{ The dataset is made available at \url{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}. This list is updated once per month---we rely on the version released on January 16, 2017.} The data from the GSA contains the following variables: (1) domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'); (2) the type of government entity to which the domain is registered, such as city, county, federal agency, etc; (3) for federal agencies, the name is specified; (4) the city in which the domain is registered. Naturally, the GSA's list does not contain cities which do not use a .gov website (or, in many cases, a city owns a registered .gov address, but uses a different one). Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa). Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the former, and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

To test whether the websites we found actually work, we use a webdriver-controlled browser [@markus, please define ``webdriver-controlled''] (Firefox/Selenium/Geckodriver). This is necessary because a) some city websites simply don't work  [@markus, why is it advantageous to use a webdriver-controlled browser for when a city website doesn't work?], and more often, b) cities sometimes change their websites' URLs, in which case they redirect from the old to the new URL. A webdriver-controlled browser, unlike the more rigid conventional scraping tools, will simply follow this redirection. This allows us to subsequently record and use the new URL for the actual website scraping. [@markus, would you add a few sentences documenting what we did to scrape the web files (i.e., used wget). Considering this section and the next one, its not clear how you get from a list of tested URLs to the website files being on your hard drive]. For some cities, whose websites make heavy use of JavaScript, this method does not lead to satisfying results. Consequently we restricted our corpus to cities with at least 3 documents.

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable, and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources. The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. The dataset is has a relative party balance, with variation in each state.

\input{tables/tabCitiesStatesParties.tex}

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is a certain degree of variance in where a city's executive authority lies. Unfortunately we do not have access to information about the type of governments across the breadth of our dataset and therefore cannot explore heterogeneity in the relationship between mayoral partisanship and municipal website contents based on the executive system variant. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. Information on other covariates (population and median household income - from the American Community Survey 5) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.

\section{The Web to Text Pipeline}

% BRUCE HERE
In this methodological pipeline from native website files to text data that is appropriate for comparative analysis, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another, but is uninformative regarding policy or process differences between governments. We describe these methodological steps in this section.

@@
